Climate Change Adaptation: Uncertainty Associated with Sequential Finite Period Decisions under Nonstationarity

### 2018-03-29 writeup/main.tex
working on discussion

\title{Adapting to Quasi-Periodic and Secular Variation of Climate Risk}
\author[1,2]{James Doss-Gollin}
\author[1,2]{David Farnham}
\author[1,2]{Upmanu Lall}
\affil[1]{Columbia Water Center, Columbia University}
\affil[2]{Department of Earth and Environmental Engineering, Columbia University}

@@

\section{Spatial Risk}

@@

@@

The assessment of sampling bias and variance associated with estimating the occurrence probability of an extreme event, $\hat{p}_T$, maps directly into this insurance contract, where a typical price for a $M$-year contract is given by
\begin{equation}\label{eq:insurance-price}
  \frac{\mathbb{C}}{M} = \lambda \qty(\mathbb{E} \qty[p_T] + c \mathbb{V} \qty[p_T])
\end{equation}
where $\mathbb{E}$ denotes the expectation operator, $\mathbb{V}$ the variance operator, $\lambda$ the profit margin, and $c$ the risk premium.
Bias in $\hat{P}_T$ leads to inefficient pricing, since if the price is too low, the insurer is exposed to failure and if it is too high the buyer will be less inclined to purchase insurance.

We illustrated sampling uncertainty and inter-model disagreement using several simple models for annual-maximum floods.
In more realistic applications, analysis of flood frequency can be improved by combining information across space \citep{Lima2016,Merz2008a}, combining model forecasts, \etc{}.
Analysis of other weather and climate hazards may use different distributions, and the time structure of the underlying climate process may be different.

For the type of extreme flood events we are interested in, the nominal return period at some location flooded during the event will typically be larger than 10 years.
While this is a relatively rare event at a given location, across a geography as large as the United States, one can expect to see such events rather frequently, emphasizing the need to understand and predict the risk of such events over a season or few years to facilitate preparation and mitigation strategies.
Recent efforts\footnote{Prad's work} to identify regions where risk of a particular event is negatively correlated in space.

Such clustering in the occurrence of extreme rainfall and drought is manifest in the potential exposure of mining companies, who face spatial as well as temporal clustering \citep{Bonnafous2017a,Bonnafous2017b}.

@@

From a decision making perspective, one could either design and build a project that gives us protection through the year 2100 considering both types of nonstationarities, or consider a sequence of projects every M years that incrementally add (or not) protection. How does the uncertainty and hence risk for these decisions manifest depending on the design level p, the duration M, and the uncertainty associated with the estimation of the \( f(Qp(t)) \) given a methodology used for prediction -- \eg{} the model chain using gcms and such and models fitted to a historical data length $M$.
What is the implication for choices of p and M under different models for stationarity and uncertainty using some simple loss and cost  functions.
If a positive discount rate is used for the future, as is typical in public and private sector planning models,\footnote{CITE} the investment returns on major public and private investment may diminish substantially far into the future.
This consideration suggests that it may be more effective to understand the consequences for analysis focusing on shorter lived mitigation alternatives that may last for 1, 5, 10, 20, 30, or 50 years.
Thus, sequences of such smaller projects may in some cases better reflect climate adaptation goals, particularly given the possibilities that major climate change impacts in a particular place may occur far into the future; that climate mitigation efforts may be successful; or that the needs of human society may change drastically over these horizons.

The approach being taken by many, including the \gls{ipcc}, is to consider discrete scenarios for what humans may do with respect to greenhouse gas emissions, and then attempt to use climate models to estimate the probability of extreme events conditional on occurrence of a particular scenario.
Because of inherent errors due to the discretization and parameterization of the equations of motion, as well as uncertainty about the underlying physical parameters, bias correction methods \citep[\eg quantile-quantile mapping;][]{Rajczak2016} are typically applied to the final outputs.
While this typically improves the model chain's representation of the historical record, this approach assumes a stationary relationship between modeled and observed rainfall at a point-by-point scale, an assumption that is unlikely to hold under the kinds of spatial and dynamical shifts anticipated under increased greenhouse gas scenarios such as expansion of the tropics and altered equator-pole temperature contrasts.\citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.\footnote{Farnham et al 2018!}

>>>>

\title{Planning for Quasi-Periodic and Secular Variability of Climate Risk}
\author[1,2]{James Doss-Gollin}
\author[1,2]{David Farnham}
\author[1,2]{Upmanu Lall}
\affil[1]{Columbia Water Center, Columbia University}
\affil[2]{Department of Earth and Environmental Engineering, Columbia University}

@@

\section{Portfolio Risk Across Space}

@@

Although the flood sequences shown in \cref{fig:example-long,fig:example-short,fig:stationary-sequences,fig:trend-sequences} exhibit the regime-like, quasi-periodic behavior observed in analysis of observational streamflow records, the assumption that the deterministic component of flood risk varies linearly with time and \gls{enso} is a simplification.
Reality, of course, is more complicated.
The probability of flooding in a particular location is modulated by multiple physical mechanisms, and dependence on a particular driver is rarely linear.
Further, the linear time dependance assumed is merely a placeholder for physical drivers, such as water storage capacity of the catchment's soil or the regional average specific humidity, that experience secular change.

More complex strategies also exist for estimating the risk of flood, or another natural hazard, from an observational record.
For example, analysis of flood frequency can be improved by combining information across space \citep{Merz2008a,Lima2016}, combining predictions from many models, or by through use of frequency- or wavelet-based methods \citep{Kwon2007}.
The approach being taken by many, including the \gls{ipcc}, is to consider discrete scenarios for what humans may do with respect to greenhouse gas emissions, and then attempt to use climate models to estimate the probability of extreme events conditional on occurrence of a particular scenario.
Because of inherent errors due to the discretization and parameterization of the equations of motion, as well as uncertainty about the underlying physical parameters, bias correction methods \citep[\eg quantile-quantile mapping;][]{Rajczak2016} are typically applied to the final outputs.
While this typically improves the model chain's representation of the historical record, this approach assumes a stationary relationship between modeled and observed rainfall at a point-by-point scale, an assumption that is unlikely to hold under the kinds of spatial and dynamical shifts anticipated under increased greenhouse gas scenarios such as expansion of the tropics and altered equator-pole temperature contrasts.\citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.\footnote{Farnham et al 2018!}

From a decision making perspective, one could either design and build a project that gives us protection through the year 2100 considering both types of nonstationarities, or consider a sequence of projects every $M$ years that incrementally add (or not) protection.
In this context one could ask how the uncertainty (and hence risk) for these decisions depends on the design level, the project duration $M$, and the uncertainty associated with the estimation of $f(\hat{p}_T)$.
This study has considered only uncertainty associated with changing flood hazard, yet the large uncertainty associated with flood risk hazard and possible adaptation technologies would necessarily be relevant.

What is the implication for choices of p and M under different models for stationarity and uncertainty using some simple loss and cost  functions.
This consideration suggests that it may be more effective to understand the consequences for analysis focusing on shorter lived mitigation alternatives that may last for 1, 5, 10, 20, 30, or 50 years.
Thus, sequences of such smaller projects may in some cases better reflect climate adaptation goals, particularly given the possibilities that major climate change impacts in a particular place may occur far into the future; that climate mitigation efforts may be successful; or that the needs of human society may change drastically over these horizons.

The results presented in \cref{sec:results} illustrate sampling uncertainty and inter-model disagreement using several simple models for annual-maximum floods.
In reality,

@@

@@

### 2018-03-29 writeup/misc.tex
reorganize writeup
working on results and intro
working on discussion

\begin{enumerate}
  \item High government indebtedness levels in many if not most emerging-market countries no longer allow public-debtd riven delivery as a scalable alternative to build urgently needed infrastructure \citep{WEF-instruments-2016}.
  \item In specific cases, investors' risk appetite and the risk level of infrastructure projects can be bridged through credit and project guarantees, insurance and other credit-enhancement schemes, also known as Risk Mitigation instruments \citep{WEF-instruments-2016}.
  \item The sustainable management of water resources requires a long-term perspective.
  Yet looking to the future reveals a host of major uncertainties, with respect to pressures from climate change, demographic change, land-use changes and other socio-economic drivers \citep{Hall2012}.
  \item Despite widespread interest in index insurance for industries such as agriculture \citep{Clarke2013} and navigation \citep{Meyer2016}, there has been little exploration of financial instruments for hydroclimate risks beyond a year or two.
  \item The insurability of risks depends on a number of factors, including:
  \begin{enumerate*}[label=(\arabic*)]
    \item mutuality (that a large pool of risk can be created);
    \item quantifiable loss;
    \item randomness of the insured event; and
    \item economic viability (that the premium is sufficient to cover losses and affordable to the policyholders).
  \end{enumerate*} \citep{Wolfrom2016}.
\end{enumerate}

>>>>

Some stuff on decisions and insurance which may be relevant for the introduction?
\begin{enumerate}
  \item High government indebtedness levels in many if not most emerging-market countries no longer allow public-debtd riven delivery as a scalable alternative to build urgently needed infrastructure \citep{WEF-instruments-2016}.
  \item In specific cases, investors' risk appetite and the risk level of infrastructure projects can be bridged through credit and project guarantees, insurance and other credit-enhancement schemes, also known as Risk Mitigation instruments \citep{WEF-instruments-2016}.
  \item The sustainable management of water resources requires a long-term perspective.
  Yet looking to the future reveals a host of major uncertainties, with respect to pressures from climate change, demographic change, land-use changes and other socio-economic drivers \citep{Hall2012}.
  \item Despite widespread interest in index insurance for industries such as agriculture \citep{Clarke2013} and navigation \citep{Meyer2016}, there has been little exploration of financial instruments for hydroclimate risks beyond a year or two.
  \item The insurability of risks depends on a number of factors, including:
  \begin{enumerate*}[label=(\arabic*)]
    \item mutuality (that a large pool of risk can be created);
    \item quantifiable loss;
    \item randomness of the insured event; and
    \item economic viability (that the premium is sufficient to cover losses and affordable to the policyholders).
  \end{enumerate*} \citep{Wolfrom2016}.
\end{enumerate}

The assessment of sampling bias and variance associated with estimating the occurrence probability of an extreme event, $\hat{p}_T$, maps directly into this insurance contract, where a typical price for a $M$-year contract is given by
\begin{equation}\label{eq:insurance-price}
  \frac{\mathbb{C}}{M} = \lambda \qty(\mathbb{E} \qty[p_T] + c \mathbb{V} \qty[p_T])
\end{equation}
where $\mathbb{E}$ denotes the expectation operator, $\mathbb{V}$ the variance operator, $\lambda$ the profit margin, and $c$ the risk premium.
Bias in $\hat{P}_T$ leads to inefficient pricing, since if the price is too low, the insurer is exposed to failure and if it is too high the buyer will be less inclined to purchase insurance.
